---
title: "Final Project Code"
author: "Helen Liang, Ivy Zhao, Xiaotong Zhao"
date: "05/10/2024"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = "/Users/hl0616/Library/Mobile Documents/com~apple~CloudDocs/NYU/Spring 2024/GPH-GU-2353 Regression I/Final Project/bike_sharing_dataset", echo = TRUE)
```

# Predicting Total Daily Rental Bike Demands

The `bike` dataset contains information such as total daily rental bike count from the Capital bikeshare system spanning from 2011 to 2012. This dataset contains 731 observations and 16 variables. However, we're only using the following 9 of those 16 variables in our study: 

* `yr`: year
  * 0 = 2011
  * 1 = 2012
* `season`: season
  * 1 = spring
  * 2 = summer
  * 3 = fall
  * 4 = winter
* `holiday`: whether the day is a holiday or not
  * 0 = no
  * 1 = yes
* `workingday`: type of day
  * 0 = no
  * 1 = yes
* `weathersit`: weather of the day
  * 1 = clear, few clouds, partly cloudy, partly cloudy
  * 2 = mist + cloudy, mist + broken clouds, mist + few clouds, mist
  * 3 = light Snow, light rain + thunderstorm + scattered clouds, light rain + scattered clouds
  * 4 = heavy rain + ice pallets + thunderstorm + mist, snow + fog
* `temp`: normalized temperature
* `atemp`: normalized feeling temperature
* `hum`: normalized humidity
* `windspeed`: normalized wind speed
* `cnt`: count of total rental bikes including both casual and registered

# Load Data
```{r}
bike <- read.csv("day.csv") #load dataset

head(bike) #view the first 6 rows
nrow(bike) #number of observations
ncol(bike) #number of variables
sum(is.na(bike)) #check for missing values
sapply(bike, class) #check for variable types
```

# Data Preparation
```{r}
library(dplyr)
library(tidyr)

bikeDf <- bike %>% 
  select(yr, season, holiday, workingday, weathersit, temp, atemp, hum, windspeed, cnt)

head(bikeDf) #view the first 6 rows
nrow(bikeDf) #number of observations
ncol(bikeDf) #number of variables
sum(is.na(bikeDf)) #check for missing values
sapply(bikeDf, class) #check for variable types
```

# Univariable Analysis
```{r}
#yr
yr_counts <- table(bikeDf$yr)
yr_percentages <- prop.table(yr_counts) * 100
yr_summary <- data.frame(yr = names(yr_counts),
Count = as.numeric(yr_counts), Percentage = yr_percentages)
print(yr_summary)

#season
season_counts <- table(bikeDf$season)
season_percentages <- prop.table(season_counts) * 100
season_summary <- data.frame(Season = names(season_counts),
Count = as.numeric(season_counts), Percentage = season_percentages)
print(season_summary)

#holiday
holiday_counts <- table(bikeDf$holiday)
holiday_percentages <- prop.table(holiday_counts) * 100
holiday_summary <- data.frame(holiday = names(holiday_counts),
Count = as.numeric(holiday_counts), Percentage = holiday_percentages)
print(holiday_summary)

#workingday
workingday_counts <- table(bikeDf$workingday)
workingday_percentages <- prop.table(workingday_counts) * 100
workingday_summary <- data.frame(workingday = names(workingday_counts),
Count = as.numeric(workingday_counts), Percentage = workingday_percentages)
print(workingday_summary)

#weathersit
weathersit_counts <- table(bikeDf$weathersit)
weathersit_percentages <- prop.table(weathersit_counts) * 100
weathersit_summary <- data.frame(weathersit = names(weathersit_counts),
Count = as.numeric(weathersit_counts), Percentage = weathersit_percentages)
print(weathersit_summary)

sd <- sapply(bikeDf[, c("cnt", "temp", "atemp", "hum", "windspeed")], sd)
print(sd)

summary(bikeDf)
```



# Density Plots
```{r}
library(ggplot2)

#temp
ggplot(bikeDf, aes(x = temp)) + 
  geom_density() +
  labs(title = "", 
       x = "Normalized Temperature",
       y = "Density")

#atemp
ggplot(bikeDf, aes(x = atemp)) +
  geom_density() +
  labs(title = "",
       x = "Normalized Feeling Temperature",
       y = "Density")

#hum
ggplot(bikeDf, aes(x = hum)) +
  geom_density() +
  labs(title = "",
       x = "Normalized Humidity",
       y = "Density")

#windspeed
ggplot(bikeDf, aes(x = windspeed)) +
  geom_density() +
  labs(title = "",
       x = "Normalized Wind Speed",
       y = "Density")

#cnt
ggplot(bikeDf, aes(x = cnt)) + 
  geom_density() +
  labs(title = "", 
       x = "Total Daily Rental Bike Count", 
       y = "Density")
```

# Initial Model (everything included)
```{r}
library(faraway)

#initial model (everything included)
initial_model <- lm(cnt ~., data = bikeDf)
summary(initial_model)

#residual vs fitted plot for initial model
plot(fitted(initial_model), rstandard(initial_model), 
     xlab = "Fitted", ylab = "Standardized Residuals")
abline(h = 0, col = "red")

#quantile-quantile plot for initial model
qqnorm(rstandard(initial_model), main = "")
abline(0, 1, col = "red")

#histogram for initial model
hist(rstandard(initial_model), main = "", xlab = "Standardized Residuals")

#########

#check for leverages
hatv <- hatvalues(initial_model)
leverages <- hatv[hatv > 2 * mean(hatv)]
halfnorm(hatv, nlab = 1, ylab = "Leverages") #595

#check for outliers
stud <- rstudent(initial_model)
stud[which.max(abs(stud))] #668
p <- 5
n <- nrow(bikeDf)
qt(1 - 0.05 / (n * 2), n - p - 1) #668 is not an outlier
outliers <- which(abs(stud) > qt(1 - 0.05 / (n * 2), n - p - 1))
outliers #668, 669

#check for cook's distances
cook <- cooks.distance(initial_model)
influential <- cook[which(cook > 0.5)]
influential #595
halfnorm(cook, nlab = 1, ylab = "Cook's Distances") #595

#remove leverages, outliers, influential points
all_indices <- unique(c(leverages, outliers, influential))
bikeDf_new <- bikeDf[-all_indices, ]
```


# Model 1 (leverages/outliers/influential points removed)
```{r}
#model1
model1 <- lm(cnt ~., data = bikeDf_new)
summary(model1)
vif(model1) #vif for model1
```

# Model 2 (with atemp removed)
```{r}
#model2 (atemp removed)
model2 <- lm(cnt ~ yr + season + holiday + workingday + weathersit + temp + hum + windspeed, data = bikeDf_new)
summary(model2)
vif(model2) #vif for model2

#residual vs fitted plot for initial model
plot(fitted(model2), rstandard(model2), 
     xlab = "Fitted", ylab = "Standardized Residuals")
abline(h = 0, col = "red")

#quantile-quantile plot for initial model
qqnorm(rstandard(model2), main = "")
abline(0, 1, col = "red")

#histogram for initial model
hist(rstandard(model2), main = "", xlab = "Standardized Residuals")

#use AIC to check if model can be improved
library(leaps)
b <- regsubsets(cnt ~ yr + season + holiday + workingday + weathersit + temp + hum + windspeed, data = bikeDf_new)

rs <- summary(b)
rs$which
a <- rs$rss
n <- nrow(bikeDf)
p <- 2:9
AIC <- n * log(a / n) + 2 * p
AIC
plot(AIC ~ I(p - 1), xlab = "Number of Predictors", ylab = "AIC") #keep model2
```

# Choosing Interaction Terms
```{r, echo = TRUE, results = "hide"}
#check for all 28 interaction term combinations
predictors <- c("yr", "season", "holiday", "workingday", "weathersit", "temp", "hum", "windspeed")
interaction_terms <- combn(predictors, 2, simplify = FALSE)
model_summaries <- list()
for(i in seq_along(interaction_terms)) {
  interaction_str <- paste(interaction_terms[[i]], collapse = "*")
  formula_str <- paste("cnt ~", paste(predictors, collapse = " + "), "+", interaction_str)
  model <- lm(as.formula(formula_str), data = bikeDf_new)
  model_summaries[[i]] <- summary(model)
  print(paste("Model", i, "with interaction", interaction_str))
  print(summary(model)$coefficients)
}
adj_r_squared <- sapply(model_summaries, function(x) x$adj.r.squared)
print(adj_r_squared)

anova_p_values <- numeric(length(interaction_terms))
for(i in seq_along(interaction_terms)) {
  interaction_str <- paste(interaction_terms[[i]], collapse = ":")
  formula_str <- paste("cnt ~", paste(predictors, collapse = " + "), "+", interaction_str)
  interaction_model <- lm(as.formula(formula_str), data = bikeDf_new)
  model_comparison <- anova(model2, interaction_model)
  anova_p_values[i] <- model_comparison$"Pr(>F)"[2]
  
  print(paste("Model with interaction", interaction_str))
  print(model_comparison)
}

names(anova_p_values) <- sapply(interaction_terms, function(x) paste(x, collapse = ":"))
print(anova_p_values)
```

# Model 3 (with 11 interaction terms)
```{r}
#significant interaction terms

#yr*season
#yr*workingday
#yr*weathersit
#yr*temp
#season*temp
#season*hum
#holiday*temp
#workingday*temp
#workingday*windspeed
#weathersit*hum
#weathersit*windspeed
#temp*windspeed

#model3
model3 <- lm(cnt ~ yr + season + holiday + workingday + weathersit + temp + hum + windspeed + yr*season + yr*workingday + yr*weathersit + yr*temp + season*temp + season*hum + holiday*temp + workingday*temp + workingday*windspeed + weathersit*hum + weathersit*windspeed + temp*windspeed, data = bikeDf_new)
summary(model3)
```

# Transformed Model 3 (with 12 interaction terms)
```{r}
library(MASS)

obj <- boxcox(model3, plotit = TRUE)
lambda <- obj$x[which.max(obj$y)]
lambda #0.7070707

#transformed model 3
trans <- (bikeDf_new$cnt^lambda)
model3.trans <- lm(trans ~ yr + season + holiday + workingday + weathersit + temp + hum + windspeed + yr*season + yr*workingday + yr*weathersit + yr*temp + season*temp + season*hum + holiday*temp + workingday*temp + workingday*windspeed + weathersit*hum + weathersit*windspeed + temp*windspeed, data = bikeDf_new)
summary(model3.trans)

#residual vs fitted plot for model3.trans
plot(fitted(model3.trans), rstandard(model3.trans), 
     xlab = "Fitted", ylab = "Standardized Residuals")
abline(h = 0, col = "red")

#quantile-quantile plot for model3.trans
qqnorm(rstandard(model3.trans), main = "")
abline(0, 1, col = "red")

#histogram for model3.trans
hist(rstandard(model3.trans), main = "", xlab = "Standardized Residuals")
```
